---
title: "Overview"
type: docs
menu:
  main:
    name: "Quickstart"
    identifier: "tutorials-advanced-quickstart-quickstart"
    parent: "tutorials-top"
    weight: 50
---

This is a short introduction to Keras advanced features. It uses:

1. `tfdatasets` to manage input data.
1. A custom model.
1. `tfautograph` for building a custom training loop.

Before running the quickstart you need to have Keras installed. Please refer to the [installation](/installation/) for installation instructions.

```{r}
library(keras)
library(tfdatasets)
library(tfautograph)
library(reticulate)
library(purrr)
```

Let's start by loading and preparing the [MNIST dataset](http://yann.lecun.com/exdb/mnist/). The values of the pixels are integers between 0 and 255, and we will convert them to floats between 0 and 1.

```{r}
mnist <- dataset_mnist()
mnist$train$x <- mnist$train$x/255
mnist$test$x <- mnist$test$x/255

dim(mnist$train$x) <- c(dim(mnist$train$x), 1)
dim(mnist$test$x) <- c(dim(mnist$test$x), 1)
```

Now let's use `tfdatasets` to batch and shuffle the dataset.

```{r}
train_ds <- mnist$train %>% 
  tensor_slices_dataset() %>%
  dataset_take(20000) %>% 
  dataset_map(~modify_at(.x, "x", tf$cast, dtype = tf$float32)) %>% 
  dataset_map(~modify_at(.x, "y", tf$cast, dtype = tf$int64)) %>% 
  dataset_shuffle(10000) %>% 
  dataset_batch(32)

test_ds <- mnist$test %>% 
  tensor_slices_dataset() %>% 
  dataset_take(2000) %>% 
  dataset_map(~modify_at(.x, "x", tf$cast, dtype = tf$float32)) %>%
  dataset_map(~modify_at(.x, "y", tf$cast, dtype = tf$int64)) %>% 
  dataset_batch(32)
```

We will now define a Keras custom model.

```{r}
simple_conv_nn <- function(filters, kernel_size) {
  keras_model_custom(name = "MyModel", function(self) {
    
    self$conv1 <- layer_conv_2d(
      filters = filters, 
      kernel_size = rep(kernel_size, 2),
      activation = "relu"
    )
    
    self$flatten <- layer_flatten()
    
    self$d1 <- layer_dense(units = 128, activation = "relu")
    self$d2 <- layer_dense(units = 10, activation = "softmax")
    
    
    function(inputs, mask = NULL) {
      inputs %>% 
        self$conv1() %>% 
        self$flatten() %>% 
        self$d1() %>% 
        self$d2()
    }
  })
}

model <- simple_conv_nn(filters = 32, kernel_size = 3)
```

We can then choose an optimizer and loss function for training:

```{r}
loss <- loss_sparse_categorical_crossentropy
optimizer <- optimizer_adam()
```

Select metrics to measure the loss and the accuracy of the model. These metrics accumulate the values over epochs and then print the overall result.

```{r}
train_loss <- tf$keras$metrics$Mean(name='train_loss')
train_accuracy <-  tf$keras$metrics$SparseCategoricalAccuracy(name='train_accuracy')

test_loss <- tf$keras$metrics$Mean(name='test_loss')
test_accuracy <- tf$keras$metrics$SparseCategoricalAccuracy(name='test_accuracy')
```

We then define a function that is able to make one training step:

```{r}
train_step <- function(images, labels) {
  
  with (tf$GradientTape() %as% tape, {
    predictions <- model(images)
    l <- loss(labels, predictions)
  })
  
  gradients <- tape$gradient(l, model$trainable_variables)
  optimizer$apply_gradients(purrr::transpose(list(
    gradients, model$trainable_variables
  )))
  
  train_loss(l)
  train_accuracy(labels, predictions)
  
}
```

We then provide a function that is able to test the model:

```{r}
test_step <- function(images, labels) {
  predictions <- model(images)
  l <- loss(labels, predictions)
  
  test_loss(l)
  test_accuracy(labels, predictions)
}
```

We can then write our training loop function:

```{r}
training_loop <- tf_function(autograph(function(train_ds, test_ds) {
  
  for (b1 in train_ds) {
    train_step(b1$x, b1$y)
  }
  
  for (b2 in test_ds) {
    test_step(b2$x, b2$y)
  }
  
  tf$print("Acc", train_accuracy$result(), "Test Acc", test_accuracy$result())
  
  train_loss$reset_states()
  train_accuracy$reset_states()
  test_loss$reset_states()
  test_accuracy$reset_states()
  
}))

```

Finally let's run our training loop for 5 epochs:

```{r, eval = FALSE}
for (epoch in 1:5) {
  cat("Epoch: ", epoch, " -----------\n")
  training_loop(train_ds, test_ds)  
}
```

```
## Epoch:  1  -----------
Acc 0.93095 Test Acc 0.954
## Epoch:  2  -----------
Acc 0.956525 Test Acc 0.95825
## Epoch:  3  -----------
Acc 0.968066692 Test Acc 0.9575
## Epoch:  4  -----------
Acc 0.9752 Test Acc 0.960125
## Epoch:  5  -----------
Acc 0.9796 Test Acc 0.9617
```





