---
title: "TensorFlow tensors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TensorFlow tensors} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
type: "docs"
menu:
  main:
    name: "Tensors"
    identifier: "custom-basic-tensors"
    parent: "custom-basic-top"
    weight: 30
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

TensorFlow, as the name indicates, is a framework to define and run computations
involving tensors. A *tensor* is a generalization of vectors and matrices to
potentially higher dimensions. Internally, TensorFlow represents tensors as
n-dimensional arrays of base datatypes.

When writing a TensorFlow program, the main object you manipulate and pass
around is the `tf$Tensor`. A `tf$Tensor` object represents a partially defined
computation that will eventually produce a value. TensorFlow programs work by
first building a graph of `tf$Tensor` objects, detailing how each tensor is
computed based on the other available tensors and then by running parts of this
graph to achieve the desired results.

A `tf$Tensor` has the following properties:

 * a data type (`float32`, `int32`, or `string`, for example)
 * a shape


Each element in the Tensor has the same data type, and the data type is always
known. The shape (that is, the number of dimensions it has and the size of each
dimension) might be only partially known. Most operations produce tensors of
fully-known shapes if the shapes of their inputs are also fully known, but in
some cases it's only possible to find the shape of a tensor at graph execution
time.

Some types of tensors are special, and these will be covered in other
units of the TensorFlow guide. The main ones are:

  * `tf$Variable`
  * `tf$constant`
  * `tf$SparseTensor`

With the exception of `tf.Variable`, the value of a tensor is immutable, which
means that in the context of a single execution tensors only have a single
value. However, evaluating the same tensor twice can return different values;
for example that tensor can be the result of reading data from disk, or
generating a random number.

## Rank

The **rank** of a `tf$Tensor` object is its number of dimensions. Synonyms for
rank include **order** or **degree** or **n-dimension**.
Note that rank in TensorFlow is not the same as matrix rank in mathematics.
As the following table shows, each rank in TensorFlow corresponds to a
different mathematical entity:

Rank | Math entity
--- | ---
0 | Scalar (magnitude only)
1 | Vector (magnitude and direction)
2 | Matrix (table of numbers)
3 | 3-Tensor (cube of numbers)
n | n-Tensor (you get the idea)


### Rank 0

The following snippet demonstrates creating a few rank 0 variables:

```{r}
library(tensorflow)
mammal <- tf$Variable("Elephant", tf$string)
ignition <- tf$Variable(451, tf$int16)
floating <- tf$Variable(3.14159265359, tf$float64)
its_complicated <- tf$Variable(12.3 - 4.85i, tf$complex64)
```

Note: A string is treated as a single object in TensorFlow, not as a sequence of
characters. It is possible to have scalar strings, vectors of strings, etc.

### Rank 1

To create a rank 1 `tf.Tensor` object, you can pass a list of items as the
initial value. For example:

```{r}
mystr <- tf$Variable(c("Hello"), tf$string)
cool_numbers <- tf$Variable(c(3.14159, 2.71828), tf$float32)
first_primes <- tf$Variable(c(2, 3, 5, 7, 11), tf$int32)
its_very_complicated <- tf$Variable(c(12.3 - 4.85i, 7.5 - 6.23i), tf$complex64)
```

### Higher ranks

A rank 2 `tf$Tensor` object consists of at least one row and at least
one column:

```{r}
mymat <- tf$Variable(list(c(7, 2),c(11, 3)), tf$int16)
```


Higher-rank Tensors, similarly, consist of an n-dimensional array. For example,
during image processing, many tensors of rank 4 are used, with dimensions
corresponding to example-in-batch, image height, image width, and color channel.

```{r}
my_image <- tf$zeros(shape = shape(10, 299, 299, 3)) # batch x height x width x color
```

### Getting a `tf$Tensor` object's rank

To determine the rank of a `tf$Tensor` object, call the `tf$rank` method.
For example, the following method programmatically determines the rank
of the `tf$Tensor` defined in the previous section:

```{r}
r <- tf$rank(my_image)
r
```

### Referring to `tf$Tensor` slices

Tensor elements can be extracted either by using functions like `tf$gather()` and `tf$slice()`, or by using `[` syntax.

Extracting tensor elements with `[` in R is similar to extracting elements from standard R arrays, albeit with some minor differences in capabilities. In contrast to most `tf$` functions, `[` defaults to R style 1-based rather than 0-based indexes. Currently, only numeric indexes in `[` are supported (no logical or character indexes)

Extracting works identically to R arrays if the slicing index is missing, supplied as a scalar, or as a sequence (e.g, created by `:` or `seq_len()`)  

```r
x[,1] )      # all rows, first column
x[1:2,] )    # first two rows, all columns
x[,1, drop = FALSE] ) # all rows, first column, but preserving the tensor rank
```

`[` also supports slices with a strided step, which can be specified in traditional R style with `seq()` or with a python-style second colon. If you are unfamiliar with python-style strided step syntax, see here for a [quick primer](https://docs.scipy.org/doc/numpy-1.13.0/reference/arrays.indexing.html#basic-slicing-and-indexing)

```r
x[, seq(1, 5, by = 2)]  # R style
x[, 1:5:2]              # Equivalent python-style strided step 
```

Missing arguments for python syntax are valid, but they must be supplied as NULL or whole expression must by backticked.

```r
x[, `::2`] 
x[, NULL:NULL:2] 
x[, `2:`] 
```

`[` also accepts  `tf$newaxis` and `all_dims()` as arguments

```r
x[,, tf$newaxis]
x[all_dims(), 1] # all_dims expands to the shape of the tensor
```

An important difference between extracting R arrays and tensorflow tensors with `[` is how negative numbers are interpreted. For tensorflow tensors, negative numbers are interpreted as selecting elements by counting from tail (e.g, they are interpreted python-style).  

```r
x[-1,]  # the last row
```

Tensors are accepted by `[` as well, but note that tensors suplied to `[` are not translated from R to python. Meaning that tensors are interpreted as 0-based, and if slicing a range with `:`, then the returned arrays is exclusive of the upper bound.

```r
x[, tf$constant(1L)] # second column
x[, tf$constant(0L):tf$constant(2L)] # first two columns
```

If you are translating existing python code to R, note that you can set an option to have all `[` arguments be interpreted pure-python style by setting `options(tensorflow.extract.style = "python")`. See ``` ?`[.tensorflow.tensor` ``` for additional options and details.

## Shape

The **shape** of a tensor is the number of elements in each dimension.
TensorFlow automatically infers shapes during graph construction. These inferred
shapes might have known or unknown rank. If the rank is known, the sizes of each
dimension might be known or unknown.

The TensorFlow documentation uses three notational conventions to describe
tensor dimensionality: rank, shape, and dimension number. The following table
shows how these relate to one another:

Rank | Shape | Dimension number | Example
--- | --- | --- | ---
0 | [] | 0-D | A 0-D tensor.  A scalar.
1 | [D0] | 1-D | A 1-D tensor with shape [5].
2 | [D0, D1] | 2-D | A 2-D tensor with shape [3, 4].
3 | [D0, D1, D2] | 3-D | A 3-D tensor with shape [1, 4, 3].
n | [D0, D1, ... Dn-1] | n-D | A tensor with shape [D0, D1, ... Dn-1].

Shapes can be represented via lists of ints, or with the
`tf$TensorShape`.

### Getting a `tf$Tensor` object's shape

There are two ways of accessing the shape of a `tf$Tensor`. While building the
graph, it is often useful to ask what is already known about a tensor's
shape. This can be done by reading the `shape` property of a `tf$Tensor` object.
This method returns a `TensorShape` object, which is a convenient way of
representing partially-specified shapes (since, when building the graph, not all
shapes will be fully known).

It is also possible to get a `tf$Tensor` that will represent the fully-defined
shape of another `tf$Tensor` at runtime. This is done by calling the `tf$shape`
operation. This way, you can build a graph that manipulates the shapes of
tensors by building other tensors that depend on the dynamic shape of the input
`tf$Tensor`.

For example, here is how to make a vector of zeros with the same size as the
number of columns in a given matrix:

```{r}
zeros <- tf$zeros(my_image$shape[2])
```

### Changing the shape of a `tf$Tensor`

The **number of elements** of a tensor is the product of the sizes of all its
shapes. The number of elements of a scalar is always `1`. Since there are often
many different shapes that have the same number of elements, it's often
convenient to be able to change the shape of a `tf.Tensor`, keeping its elements
fixed. This can be done with `tf.reshape`.

The following examples demonstrate how to reshape tensors:

```{r}
rank_three_tensor <- tf$ones(shape(3, 4, 5))
matrix <- tf$reshape(rank_three_tensor, shape(6, 10))  # Reshape existing content into
                                                 # a 6x10 matrix
matrixB <- tf$reshape(matrix, shape(3, -1))  #  Reshape existing content into a 3x20
                                       # matrix. -1 tells reshape to calculate
                                       # the size of this dimension.
matrixAlt <- tf$reshape(matrixB, shape(4, 3, -1))  # Reshape existing content into a
                                             #4x3x5 tensor

# Note that the number of elements of the reshaped Tensors has to match the
# original number of elements. Therefore, the following example generates an
# error because no possible value for the last dimension will match the number
# of elements.
```

## Data types

In addition to dimensionality, Tensors have a data type. Refer to the
`tf$DType` page for a complete list of the data types.

It is not possible to have a `tf$Tensor` with more than one data type. It is
possible, however, to serialize arbitrary data structures as `string`s and store
those in `tf$Tensor`s.

It is possible to cast `tf$Tensor`s from one datatype to another using
`tf$cast`:

``` {r}
# Cast a constant integer tensor into floating point.
float_tensor <- tf$cast(tf$constant(c(1L, 2L, 3L)), dtype=tf$float32)
```

To inspect a `tf$Tensor`'s data type use the `Tensor.dtype` property.

When creating a `tf$Tensor` from an R object you may optionally specify the
datatype. If you don't, TensorFlow chooses a datatype that can represent your
data. TensorFlow converts R integers to `tf$int32` and R floating
point numbers to `tf$float32`. Otherwise TensorFlow uses the same rules numpy
uses when converting to arrays.



