---
title: "TensorFlow variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TensorFlow variables} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
type: "docs"
menu:
  main:
    name: "Variables"
    identifier: "custom-basic-variables"
    parent: "custom-basic-top"
    weight: 20
aliases:
  - /tensorflow/articles/howto_variables.html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

A TensorFlow **variable** is the best way to represent shared, persistent state
manipulated by your program.

Variables are manipulated via the `tf$Variable` class. A `tf$Variable`
represents a tensor whose value can be changed by running ops on it.  Specific
ops allow you to read and modify the values of this tensor. Higher level
libraries like Keras use `tf$Variable` to store model parameters. This
guide covers how to create, update, and manage `tf$Variable`s in TensorFlow.

## Create a variable

To create a variable, simply provide the initial value

```{r}
library(tensorflow)
my_variable <- tf$Variable(tf$zeros(shape(1,2,3)))
```

This creates a variable which is a three-dimensional tensor with shape `[1, 2,
3]` filled with zeros. This variable will, by default, have the `dtype`
`tf$float32`. The dtype is, if not specified, inferred from the initial
value.

If there's a `tf$device` scope active, the variable will be placed on that
device; otherwise the variable will be placed on the "fastest" device compatible
with its dtype (this means most variables are automatically placed on a GPU if
one is available). For example, the following snippet creates a variable named
`v` and places it on the second GPU device:

```{r, eval = FALSE}
with(tf$device("/device:GPU:1"), {
  v <- tf$Variable(tf$zeros(c(10, 10)))
})
```

Ideally though you should use the `tf$distribute` API, as that allows you to
write your code once and have it work under many different distributed setups.

## Use a variable

To use the value of a `tf$Variable` in a TensorFlow graph, simply treat it like
a normal `tf$Tensor`:

```{r}
v <- tf$Variable(0)
w <- v + 1 # w is a tf.Tensor which is computed based on the value of v.
# Any time a variable is used in an expression it gets automatically
# converted to a tf$Tensor representing its value.
```

To assign a value to a variable, use the methods `assign`, `assign_add`, and
friends in the `tf$Variable` class. For example, here is how you can call these
methods:

```{r}
v <- tf$Variable(0)
v$assign_add(1)
```

Most TensorFlow optimizers have specialized ops that efficiently update the
values of variables according to some gradient descent-like algorithm. See
`tf$keras$optimizers$Optimizer` for an explanation of how to use optimizers.

You can also explicitly read the current value of a variable, using
`read_value`:

```{r}
v <- tf$Variable(0)
v$assign_add(1)
v$read_value()
```

When the last reference to a `tf$Variable` goes out of scope its memory is
freed.

