---
title: "Transfer learning with tfhub"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: Text Classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/keras
menu:
  main:
    name: "Transfer learning with tfhub"
    identifier: "keras-tutorial-text-classification-tfhub"
    parent: "tutorials-beginners-basic-ml-top"
    weight: 35
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

This tutorial classifies movie reviews as positive or negative using the text of the review. This is an example of binary — or two-class — classification, an important and widely applicable kind of machine learning problem.

We'll use the [IMDB dataset](https://keras.rstudio.com/reference/dataset_imdb.html) that contains the text of 50,000 movie reviews from the [Internet Movie Database](https://www.imdb.com/). These are split into 25,000 reviews for training and 25,000 reviews for testing. The training and testing sets are balanced, meaning they contain an equal number of positive and negative reviews.

We will use [Keras](https://github.com/rstudio/keras) to build and train the model
and [tfhub](https://github.com/rstudio/tfhub) for Transfer Learning. We will also use [tfds](https://github.com/rstudio/tfds) to load the IMDB dataset.

Let's start and load the required libraries.

```{r message=FALSE, warning=FALSE}
library(keras)
library(tfhub)
library(tfds)
library(tfdatasets)
```

## Download the IMDB dataset

The IMDB dataset is available on [imdb reviews](https://github.com/tensorflow/datasets/blob/master/docs/datasets.md#imdb_reviews) or on [tfds](https://github.com/rstudio/tfds). The one that comes packaged with Keras is already pre-processed so it's not useful for this tutorial.

The following code downloads the IMDB dataset to your machine:

```{r}
imdb <- tfds_load(
  "imdb_reviews:1.0.0", 
  split = list("train[:60%]", "train[-40%:]", "test"), 
  as_supervised = TRUE
)
summary(imdb)
```

`tfds_load` returns a TensorFlow Dataset, an abstraction that represents a sequence of elements, in which each element consists of one or more components. 

To access individual elements of a dataset you can use:

```{r}
first <- imdb[[1]] %>% 
  dataset_batch(1) %>% # Used to get only the first example
  reticulate::as_iterator() %>% 
  reticulate::iter_next()
str(first)
```

We will see next that Keras knows how to extract elements from TensorFlow Datasets
automatically, making it a much more memory efficient alternative than loading the entire dataset into RAM before passing it to Keras.

## Build the model

The neural network is created by stacking layers — this requires three main architectural decisions:

1. How to represent the text?
1. How many layers to use in the model?
1. How many hidden units to use for each layer?

In this example, the input data consists of sentences. The labels to predict are either 0 or 1.

One way to represent the text is to convert sentences into embedding vectors. We can use a pre-trained text embedding as the first layer, which will have three advantages: * we don't have to worry about text preprocessing, * we can benefit from transfer learning, * the embedding has a fixed size, so it's simpler to process.

For this example we will use a pre-trained text embedding model from [TensorFlow Hub](https://github.com/rstudio/tfhub) called [google/tf2-preview/gnews-swivel-20dim/1](https://tfhub.dev/google/tf2-preview/gnews-swivel-20dim/1).

There are three other pre-trained models to test for the sake of this tutorial:

* [google/tf2-preview/gnews-swivel-20dim-with-oov/1](https://tfhub.dev/google/tf2-preview/gnews-swivel-20dim-with-oov/1) - same as google/tf2-preview/gnews-swivel-20dim/1, but with 2.5% vocabulary converted to OOV buckets. This can help if the vocabulary of the task and the vocabulary of the model don't fully overlap.
* [google/tf2-preview/nnlm-en-dim50/1](https://tfhub.dev/google/tf2-preview/nnlm-en-dim50/1) - A much larger model with ~1M vocabulary size and 50 dimensions.
* [google/tf2-preview/nnlm-en-dim128/1](https://tfhub.dev/google/tf2-preview/nnlm-en-dim128/1) - Even larger model with ~1M vocabulary size and 128 dimensions.

Let's first create a Keras layer that uses a TensorFlow Hub model to embed the sentences, and try it out on a couple of input examples. Note that no matter the length of the input text, the output shape of the embeddings is: `(num_examples, embedding_dimension)`.

```{r}
embedding_layer <- layer_hub(handle = "https://tfhub.dev/google/tf2-preview/gnews-swivel-20dim/1")
embedding_layer(first[[1]])
```

Let's now build the full model:

```{r}
model <- keras_model_sequential() %>% 
  layer_hub(
    handle = "https://tfhub.dev/google/tf2-preview/gnews-swivel-20dim/1",
    input_shape = list(),
    dtype = tf$string,
    trainable = TRUE
  ) %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

summary(model)
```

The layers are stacked sequentially to build the classifier:

1. The first layer is a TensorFlow Hub layer. This layer uses a pre-trained Saved Model to map a sentence into its embedding vector. The pre-trained text embedding model that we are using ([google/tf2-preview/gnews-swivel-20dim/1](https://tfhub.dev/google/tf2-preview/gnews-swivel-20dim/1)) splits the sentence into tokens, embeds each token and then combines the embedding. The resulting dimensions are: (num_examples, embedding_dimension).
1. This fixed-length output vector is piped through a fully-connected (dense) layer with 16 hidden units.
1. The last layer is densely connected with a single output node. Using the sigmoid activation function, this value is a float between 0 and 1, representing a probability, or confidence level.

Let's compile the model.

## Loss function and optimizer

A model needs a loss function and an optimizer for training. Since this is a binary classification problem and the model outputs a probability (a single-unit layer with a sigmoid activation), we'll use the `binary_crossentropy` loss function.

This isn't the only choice for a loss function, you could, for instance, choose mean_squared_error. But, generally, `binary_crossentropy` is better for dealing with probabilities — it measures the "distance" between probability distributions, or in our case, between the ground-truth distribution and the predictions.

Now, configure the model to use an optimizer and a loss function:

```{r}
model %>% 
  compile(
    optimizer = "adam",
    loss = "binary_crossentropy",
    metrics = "accuracy"
  )
```

## Train the model

Train the model for 20 epochs in mini-batches of 512 samples. This is 20 iterations over all samples in the dataset. While training, monitor the model's loss and accuracy on the 10,000 samples from the validation set:

```{r}
model %>% 
  fit(
    imdb[[1]] %>% dataset_shuffle(10000) %>% dataset_batch(512),
    epochs = 20,
    validation_data = imdb[[2]] %>% dataset_batch(512),
    verbose = 2
  )
```

## Evaluate the model

And let's see how the model performs. Two values will be returned. Loss (a number which represents our error, lower values are better), and accuracy.

```{r}
model %>% 
  evaluate(imdb[[3]] %>% dataset_batch(512), verbose = 0)
```

This fairly naive approach achieves an accuracy of about 87%. With more advanced approaches, the model should get closer to 95%.