---
title: "Installing TensorFlow"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

Prior to using the **tensorflow** R package you need to install a
version of TensorFlow on your system. Below we describe how to install
TensorFlow as well the various options available for customizing your
installation.

Note that this article principally covers the use of the R
`install_tensorflow()` function, which provides an easy to use wrapper
for the various steps required to install TensorFlow.

You can also choose to install TensorFlow manually (as described at
<https://www.tensorflow.org/install/>). In that case the [Custom
Installation] section covers how to arrange for the tensorflow R package
to use the version you installed.

## Quick Start

First, install the tensorflow R package from GitHub as follows:

```{r}
devtools::install_github("rstudio/tensorflow")
```

Then, use the `install_tensorflow()` function to install TensorFlow:

```{r}
library(tensorflow)
install_tensorflow()
```

You can confirm that the installation succeeded with:

```{r}
sess = tf$Session()
hello <- tf$constant('Hello, TensorFlow!')
sess$run(hello)
```

This will provide you with a default installation of TensorFlow suitable
for use with the tensorflow R package. Read on if you want to learn
about additional installation options, including installing a version of
TensorFlow that takes advantage of Nvidia GPUs if you have the correct
CUDA libraries installed.

## Installation Methods

TensorFlow is distributed as a Python package and so needs to be
installed within a Python environment on your system. By default, the
`install_tensorflow()` function attempts to install TensorFlow within an
isolated Python environment ("r-reticulate").

These are the available methods and their behavior:

| Method     | Description                                                                |
|------------|----------------------------------------------------------------------------|
| auto       | Automatically choose an appropriate default for the current platform.      |
| virtualenv | Install into a Python virtual environment at `~/.virtualenvs/r-reticulate` |
| conda      | Install into an Anaconda Python environment named `r-reticulate`           |
| system     | Install into the system Python environment                                 |

The "virtualenv" and "conda" methods are available on Linux and OS X and
the "conda" method is available on Windows.

Note that the "virtualenv" method uses the system version of Python
(i.e. one located at either `/usr/bin` or `/usr/local/bin`, in that
order) rather than other versions of Python that may be installed. This
is to ensure that the virtual environment is created with a version and
configuration of Python that is widely tested with TensorFlow. If you
want to install TensorFlow within another version of Python see the
[Custom Installation] section below.

To override the default behavior use the `method` parameter, for
example:

```{r}
install_tensorflow(method = "conda")
```

You can also install TensorFlow within a different environment name, in
this case you can use the `envname` argument:

```{r}
install_tensorflow(envname = "r-tensorflow")
```

## CPU vs. GPU

TensorFlow can be configured to run on either CPUs or GPUs. The CPU
version is much easier to install and configure so is the best starting
place especially when you are first learning how to use TensorFlow.
Here's the guidance on CPU vs. GPU versions from the TensorFlow website:

-   **TensorFlow with CPU support only**. If your system does not have a
    NVIDIA® GPU, you must install this version. Note that this version
    of TensorFlow is typically much easier to install (typically, in 5
    or 10 minutes), so even if you have an NVIDIA GPU, we recommend
    installing this version first.

-   **TensorFlow with GPU support**. TensorFlow programs typically run
    significantly faster on a GPU than on a CPU. Therefore, if your
    system has a NVIDIA® GPU meeting the prerequisites shown below and
    you need to run performance-critical applications, you should
    ultimately install this version.

By default `install_tensorflow()` installs the CPU version of
TensorFlow.

If you want to install the GPU version please see the article on
[installing the GPU
version](https://tensorflow.rstudio.com/tools/local_gpu.html).

## Alternate Versions

By default, `install_tensorflow()` install the latest release version of
TensorFlow. You can override this behavior by specifying the `version`
parameter. For example:

```{r}
install_tensorflow(version = "1.0.0")
```

Note that this should be a full `major.minor.patch` version
specification (rather than just major and minor versions).

You can install the nightly build of TensorFlow (CPU or GPU version)
with:

```{r}
install_tensorflow(version = "nightly")      # cpu version
install_tensorflow(version = "nightly-gpu")  # gpu version
```

You can install any other build of TensorFlow by specifying a URL to a
TensorFlow binary. For example:

```{r}
install_tensorflow(version = "https://files.pythonhosted.org/packages/c2/c1/a035e377cf5a5b90eff27f096448fa5c5a90cbcf13b7eb0673df888f2c2d/tf_nightly-1.12.0.dev20180918-cp36-cp36m-manylinux1_x86_64.whl")
```

Nightly builds for different OSes and architectures are linked on the
[TensorFlow repository root
page](https://github.com/tensorflow/tensorflow).

## Custom Installation

The `install_tensorflow()` function is provided as a convenient way to
get started, but is not required. If you have an existing installation
of TensorFlow or just prefer your own custom installation that's fine
too.

The full instructions for installing TensorFlow on various platforms are
here: <https://www.tensorflow.org/install/>. After installing, please
refer to the sections below on locating TensorFlow and meeting
additional dependencies to ensure that the tensorflow for R package
functions correctly with your installation.

### Locating TensorFlow

Once you've installed TensorFlow you need to ensure that the tensorflow
for R package can find your installation. The package scans the system
for various versions of Python, and also scans available virtual
environments and conda environments, so in many cases things will just
work without additional effort.

If the version of TensorFlow you installed is not found automatically,
then you can use the following techniques to ensure that TensorFlow is
located.

Specify the `TENSORFLOW_PYTHON` environment variable to force probing
within a specific Python installation. For example:

```{r}
library(tensorflow)
Sys.setenv(TENSORFLOW_PYTHON="/usr/local/bin/python")
```

Alternatively, call the `use_python` family of configuration functions:

| Function           | Description                                           |
|--------------------|-------------------------------------------------------|
| `use_python()`     | Specify the path a specific Python binary.            |
| `use_virtualenv()` | Specify the directory containing a Python virtualenv. |
| `use_condaenv()`   | Specify the name of a conda environment.              |

Note that `use_virtualenv` is not supported on Windows.

For example:

```{r}
library(tensorflow)
use_python("/usr/local/bin/python")
use_virtualenv("~/myenv")
use_condaenv("myenv")
```

Note that you can include multiple calls to the `use_` functions and all
provided locations will be tried in the order they were specified.

You can also use the `required` argument from the `use_*` functions. In
this case TensorFlow will only be searched at the specified location.

```{r}
library(tensorflow)
use_virtualenv("my_env", required = TRUE)
```

### Additional Dependencies

There are some components of TensorFlow (e.g. the
[Keras](https://github.com/tensorflow/tensorflow/tree/master/tensorflow/contrib/keras)
library) which have dependencies on additional Python packages. The
`install_tensorflow()` function installs these dependencies
automatically, however if you do a custom installation you should be
sure to install them manually.

You can install the additional dependencies with the following command:

``` bash
pip install h5py pyyaml requests Pillow scipy
```

## Supported Platforms

Note that binary installations of TensorFlow are provided for Windows,
OS X, and Ubuntu 14.04 or higher. It's possible that binary
installations will work on other Linux variants but Ubuntu is the only
platform tested and supported.

In particular, if you are running on RedHat or CentOS you will need to
[install from
source](https://www.tensorflow.org/install/install_sources) then follow
the instructions in the [Custom Installation] section to ensure that
your installation of TensorFlow can be used with the tensorflow R
package.

