---
title: "Classify structured data with feature columns"
type: docs
menu:
  main:
    name: "Classify structured data"
    identifier: "tutorials-advanced-structured-classify"
    parent: "tutorials-advanced-structured-top"
    weight: 20
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

This tutorial demonstrates how to classify structured data (e.g. tabular data in a CSV). We will use Keras to define the model, and feature columns as a bridge to map from columns in a CSV to features used to train the model. This tutorial contains complete code to:

1. Load a CSV file using the tidyverse.
2. Build an input pipeline to batch and shuffle the rows using tf.data.
3. Map from columns in the CSV to features used to train the model using feature columns.
4. Build, train, and evaluate a model using Keras.

## The Dataset

We will use a small dataset provided by the Cleveland Clinic Foundation for Heart Disease. There are several hundred rows in the CSV. Each row describes a patient, and each column describes an attribute. We will use this information to predict whether a patient has heart disease, which in this dataset is a binary classification task.

Following is a description of this dataset. Notice there are both numeric and categorical columns.

>Column| Description| Feature Type | Data Type
>------------|--------------------|----------------------|-----------------
>Age | Age in years | Numerical | integer
>Sex | (1 = male; 0 = female) | Categorical | integer
>CP | Chest pain type (0, 1, 2, 3, 4) | Categorical | integer
>Trestbpd | Resting blood pressure (in mm Hg on admission to the hospital) | Numerical | integer
>Chol | Serum cholestoral in mg/dl | Numerical | integer
>FBS | (fasting blood sugar > 120 mg/dl) (1 = true; 0 = false) | Categorical | integer
>RestECG | Resting electrocardiographic results (0, 1, 2) | Categorical | integer
>Thalach | Maximum heart rate achieved | Numerical | integer
>Exang | Exercise induced angina (1 = yes; 0 = no) | Categorical | integer
>Oldpeak | ST depression induced by exercise relative to rest | Numerical | integer
>Slope | The slope of the peak exercise ST segment | Numerical | float
>CA | Number of major vessels (0-3) colored by flourosopy | Numerical | integer
>Thal | 3 = normal; 6 = fixed defect; 7 = reversable defect | Categorical | string
>Target | Diagnosis of heart disease (1 = true; 0 = false) | Classification | integer

## Setup

We will use Keras and TensorFlow datasets.

```{r}
library(keras)
library(tfdatasets)
library(tidyverse)
library(rsample)
```

## Read the data

We will use `read_csv` in order to read the csv file to R. 

```{r}
heart <- pins::pin("https://storage.googleapis.com/applied-dl/heart.csv", "heart")
df <- read_csv(heart)
glimpse(df)
```

## Split the dataframe into train, validation, and test

We are going to use the `rsample` package to split the data into train, validation
and test sets.

```{r}
# first we split between training and testing sets
split <- initial_split(df, prop = 4/5)
train <- training(split)
test <- testing(split)

# the we split the training set into validation and training
split <- initial_split(train, prop = 4/5)
train <- training(split)
val <- testing(split)
```

```{r}
nrow(train)
nrow(val)
nrow(test)
```

## Create an input pipeline using tfdatasets

Next, we will wrap the dataframes with tfdatasets. This will enable us to use feature columns as a bridge to map from the columns in the dataframe to features used to train the model. If we were working with a very large CSV file (so large that it does not fit into memory), we would use tfdatasets to read it from disk directly. That is not covered in this tutorial.

```{r}
df_to_dataset <- function(df, shuffle = TRUE, batch_size = 32) {
  ds <- df %>% 
    tensor_slices_dataset()
  
  if (shuffle)
    ds <- ds %>% dataset_shuffle(buffer_size = nrow(df))
  
  ds %>% 
    dataset_batch(batch_size = batch_size)
}
```

```{r}
batch_size <- 5
train_ds <- df_to_dataset(train, batch_size = batch_size)
val_ds <- df_to_dataset(val, shuffle = FALSE, batch_size = batch_size)
test_ds <- df_to_dataset(test, shuffle = FALSE, batch_size = batch_size)
```

## Understand the input pipeline

Now that we have created the input pipeline, let's call it to see the format of the data it returns. We have used a small batch size to keep the output readable.

```{r}
train_ds %>% 
  reticulate::as_iterator() %>% 
  reticulate::iter_next() %>% 
  str()
```

We can see that the dataset returns a list of column names (from the dataframe) that map to column values from rows in the dataframe.

## Create the feature spec

We want to train a model to predict the `target` variable using Keras but, before
that we need to prepare the data. We need to transform the categorical variables
into some form of dense variable, we usually want to normalize all numeric columns too.

The feature spec interface works with `data.frame`s or TensorFlow datasets objects.

Let's start creating our feature specification:

```{r}
spec <- feature_spec(train_ds, target ~ .)
```

The first thing we need to do after creating the feature_spec is decide on the variables' types.

We can do this by adding steps to the `spec` object.

```{r}
spec <- spec %>% 
  step_numeric_column(
    all_numeric(), -cp, -restecg, -exang, -sex, -fbs,
    normalizer_fn = scaler_standard()
  ) %>% 
  step_categorical_column_with_vocabulary_list(thal)
```

The following steps can be used to define the variable type:

* `step_numeric_column` to define numeric variables
* `step_categorical_with_vocabulary_list` for categorical variables with a fixed vocabulary
* `step_categorical_column_with_hash_bucket` for categorical variables using the hash trick
* `step_categorical_column_with_identity` to store categorical variables as integers
* `step_categorical_column_with_vocabulary_file` when you have the possible vocabulary in a file

When using `step_categorical_column_with_vocabulary_list` you can also provide a `vocabulary` argument
with the fixed vocabulary. The recipe will find all the unique values in the dataset and use it
as the vocabulary.

You can also specify a `normalizer_fn` to the `step_numeric_column`. In this case the variable will be
transformed by the feature column. Note that the transformation will occur in the TensorFlow Graph,
so it must use only TensorFlow ops. Like in the example we offer pre-made normalizers - and they will
compute the normalizing function during the recipe preparation.

You can also use selectors like:

* `starts_with()`, `ends_with()`, `matches()` etc. (from tidyselect)
* `all_numeric()` to select all numeric variables
* `all_nominal()` to select all strings
* `has_type("float32")` to select based on TensorFlow variable type.

Now we can print the recipe:

```{r}
spec
```

After specifying the types of the columns you can add transformation steps. 
For example you may want to bucketize a numeric column:

```{r}
spec <- spec %>% 
  step_bucketized_column(age, boundaries = c(18, 25, 30, 35, 40, 45, 50, 55, 60, 65))
```

You can also specify the kind of numeric representation that you want to use for
your categorical variables.

```{r}
spec <- spec %>% 
  step_indicator_column(thal) %>% 
  step_embedding_column(thal, dimension = 2)
```

Another common transformation is to add interactions between variables using crossed
columns. 

```{r}
spec <- spec %>% 
  step_crossed_column(thal_and_age = c(thal, bucketized_age), hash_bucket_size = 1000) %>% 
  step_indicator_column(thal_and_age)
```

Note that the `crossed_column` is a categorical column, so we need to also specify what
kind of numeric tranformation we want to use. Also note that we can name the transformed
variables - each step uses a default naming for columns, eg. `bucketized_age` is the
default name when you use `step_bucketized_column` with column called `age`.

With the above code we have created our recipe. Note we can also define the
recipe by chaining a sequence of methods:

```{r} 
spec <- feature_spec(train_ds, target ~ .) %>% 
  step_numeric_column(
    all_numeric(), -cp, -restecg, -exang, -sex, -fbs,
    normalizer_fn = scaler_standard()
  ) %>% 
  step_categorical_column_with_vocabulary_list(thal) %>% 
  step_bucketized_column(age, boundaries = c(18, 25, 30, 35, 40, 45, 50, 55, 60, 65)) %>% 
  step_indicator_column(thal) %>% 
  step_embedding_column(thal, dimension = 2) %>% 
  step_crossed_column(c(thal, bucketized_age), hash_bucket_size = 10) %>%
  step_indicator_column(crossed_thal_bucketized_age)
```

After defining the recipe we need to `fit` it. It's when fitting that we compute the vocabulary
list for categorical variables or find the mean and standard deviation for the normalizing functions.
Fitting involves evaluating the full dataset, so if you have provided the vocabulary list and 
your columns are already normalized you can skip the fitting step (TODO).

In our case, we will fit the feature spec, since we didn't specify the vocabulary list
for the categorical variables.

```{r}
spec_prep <- fit(spec)
```

After preparing we can see the list of dense features that were defined:

```{r}
str(spec_prep$dense_features())
```

## Build the model

Now we are ready to define our model in Keras. We will use a specialized `layer_dense_features` that
knows what to do with the feature columns specification.

We also use a new `layer_input_from_dataset` that is useful to create a Keras input object copying the structure from a `data.frame` or TensorFlow dataset.

```{r}
model <- keras_model_sequential() %>% 
  layer_dense_features(dense_features(spec_prep)) %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")


model %>% compile(
  loss = loss_binary_crossentropy, 
  optimizer = "adam", 
  metrics = "binary_accuracy"
)
```

## Train the model

We can finally train the model on the dataset:

```{r warning=FALSE}
history <- model %>% 
  fit(
    dataset_use_spec(train_ds, spec = spec_prep),
    epochs = 15, 
    validation_data = dataset_use_spec(val_ds, spec_prep),
    verbose = 2
  )
plot(history)
```

Finally we can make predictions in the test set and calculate performance 
metrics like the AUC of the ROC curve:

```{r}
pred <- predict(model, test)
Metrics::auc(test$target, pred)
```

