---
title: "Imbalanced classification: credit card fraud detection"
authors: 
  - "[fchollet](https://twitter.com/fchollet)"
  - "[terrytangyuan](https://github.com/terrytangyuan) - R translation"
date-created: 2022/11/22
date-last-modified: 2022/11/22
description: Demonstration of how to handle highly imbalanced classification problems.
categories: [structured]
aliases: 
  - ../guide/keras/examples/imbalanced_classification/index.html
---

## Introduction

This example looks at the
[Kaggle Credit Card Fraud Detection](https://www.kaggle.com/mlg-ulb/creditcardfraud/)
dataset to demonstrate how
to train a classification model on data with highly imbalanced classes.


## First, vectorize the CSV data

```{r}
library(tensorflow)
library(keras)

# Get the real data from https://www.kaggle.com/mlg-ulb/creditcardfraud/
df <- read.csv2("/Users/yuan.tang/Downloads/creditcard.csv", header = TRUE, sep = ",")

sprintf("HEADER: %s", paste0(colnames(df), collapse = ","))
sprintf("EXAMPLE FEATURES: %s", paste0(df[1, ], collapse = ","))
features <- df[,-ncol(df)]
targets <- df[ncol(df)]
sprintf("features.shape: %s", paste0(dim(features), collapse = ", "))
sprintf("targets.shape: %s", paste0(dim(targets), collapse = ","))
```

## Prepare a validation set

```{r}
num_val_samples <- sample(nrow(df), ceiling(nrow(df)*0.2))
train_features <- features[-num_val_samples,]
train_targets <- as.data.frame(targets[-num_val_samples,])
val_features <- features[num_val_samples,]
val_targets <- as.data.frame(targets[num_val_samples,])

sprintf("Number of training samples: %s", nrow(train_features))
sprintf("Number of validation samples: %s", nrow(val_features))
```

## Analyze class imbalance in the targets

```{r}
counts <- table(train_targets)
sprintf("Number of positive samples in training data: %s (%.2f%% of total)", counts[["1"]], 100 * (counts[["1"]] / nrow(train_features)))
```

```{r}
weight_for_0 = 1.0 / counts[["0"]]
weight_for_1 = 1.0 / counts[["1"]]
```

## Normalize the data using training set statistics

```{r}
means <- apply(train_features, 2, function(x) {
  mean(as.numeric(x))
})
stds <- apply(train_features, 2, function(x) {
  sd(as.numeric(x))
})
for (i in ncol(train_features)) {
  train_features[, i] <- (as.numeric(train_features[, i]) - means[i]) / stds[i]
  val_features[, i] <- (as.numeric(val_features[, i]) - means[i]) / stds[i]
}
```

## Build a binary classification model

```{r}
model <- keras_model_sequential() %>%
  layer_dense(256, activation = "relu", input_shape = tuple(dim(train_features)[-1])) %>%
  layer_dense(256, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(256, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(1, activation = "sigmoid")

summary(model)
```

## Train the model with `class_weight` argument

```{r}
metrics <- list(
  metric_false_negatives(name = "fn"),
  metric_false_positives(name = "fp"),
  metric_true_negatives(name = "tn"),
  metric_true_positives(name = "tp"),
  metric_precision(name = "precision"),
  metric_recall(name = "recall")
)
model %>% compile(
  optimizer = optimizer_adam(1e-2),
  loss = "binary_crossentropy",
  metrics = metrics
)
class_weight <- list("0" = weight_for_0, "1" = weight_for_1)
callbacks <- c(callback_model_checkpoint("fraud_model_at_epoch_{epoch}.h5"))

train_features <- as.matrix(apply(train_features, 2, as.numeric))
train_targets <- as.matrix(apply(train_targets, 2, as.numeric))
validation_data <- list(as.matrix(apply(val_features, 2, as.numeric)), as.matrix(apply(val_targets, 2, as.numeric)))

model %>% fit(
  train_features, train_targets, batch_size = 2048,
  epochs = 30, verbose = 2, validation_data = validation_data,
  callbacks = callbacks, class_weight = class_weight)
```

## Conclusions

At the end of training, out of 56,961 validation transactions, we are:

- Correctly identifying 66 of them as fraudulent
- Missing 9 fraudulent transactions
- At the cost of incorrectly flagging 441 legitimate transactions

In the real world, one would put an even higher weight on class 1,
so as to reflect that False Negatives are more costly than False Positives.

Next time your credit card gets  declined in an online purchase -- this is why.

Example available on HuggingFace.
| Trained Model | Demo |
| :--: | :--: |
| [![Generic badge](https://img.shields.io/badge/🤗%20Model-Imbalanced%20Classification-black.svg)](https://huggingface.co/keras-io/imbalanced_classification) | [![Generic badge](https://img.shields.io/badge/🤗%20Spaces-Imbalanced%20Classification-black.svg)](https://huggingface.co/spaces/keras-io/Credit_Card_Fraud_Detection) |
