---
title: "Custom Installation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

The `install_tensorflow()` function is provided as a convenient way to
get started, but is not required. If you have an existing installation
of TensorFlow or just prefer your own custom installation that's fine
too.

The full instructions for installing TensorFlow on various platforms are
here: <https://www.tensorflow.org/install/>. After installing, please
refer to the sections below on locating TensorFlow and meeting
additional dependencies to ensure that the tensorflow for R package
functions correctly with your installation.

### Locating TensorFlow

Once you've installed TensorFlow you need to ensure that the tensorflow
for R package can find your installation. The package scans the system
for various versions of Python, and also scans available virtual
environments and conda environments, so in many cases things will just
work without additional effort.

If the version of TensorFlow you installed is not found automatically,
then you can use the following techniques to ensure that TensorFlow is
located.

Specify the `RETICULATE_PYTHON` environment variable to force probing
within a specific Python installation. For example:

```{r}
Sys.setenv(RETICULATE_PYTHON="/usr/local/bin/python")
library(tensorflow)
```

You could also add the `RETICULATE_PYTHON` environment variable to your
`.RProfile` or `.REnviron`.

Alternatively, call the `use_python()` family of configuration
functions:

| Function           | Description                                           |
|--------------------|-------------------------------------------------------|
| `use_python()`     | Specify the path a specific Python binary.            |
| `use_virtualenv()` | Specify the directory containing a Python virtualenv. |
| `use_condaenv()`   | Specify the name of a conda environment.              |

For example:

```{r}
library(tensorflow)
use_virtualenv("my_env")
```

Note that if you use the argument `required = FALSE`, you can include
multiple calls to the `use_` functions and all provided locations will
be tried in the order they were specified. For example:

```{r}
library(tensorflow)
use_python("/usr/local/bin/python", required = FALSE)
use_virtualenv("~/myenv", required = TRUE)
use_condaenv("myenv", required = TRUE)
```

### Additional Dependencies

There are some components of TensorFlow (e.g. the
[Keras](https://github.com/tensorflow/tensorflow/tree/master/tensorflow/python/keras)
library) which have dependencies on additional Python packages. The
`install_keras()` function installs these dependencies automatically,
however if you do a custom installation you should be sure to install
them manually.

You can install the additional dependencies with the following command:

```{r, eval = TRUE, results = 'asis', echo = FALSE}
cat(sep = "", "\n``` bash
pip install ", paste(keras:::default_extra_packages(), collapse = " "), 
"\n```")
```

## Supported Platforms

Note that binary installations of TensorFlow are provided for Windows,
OS X, and Ubuntu 16.04 or higher. It's possible that binary
installations will work on other Linux variants but Ubuntu is the only
platform tested and supported.

In particular, if you are running on RedHat or CentOS you may need to
[install from source](https://www.tensorflow.org/install/source) then
follow the instructions in the Custom Installation section to ensure
that your installation of TensorFlow can be used with the tensorflow R
package.
