---
title: "Overview"
type: docs
menu:
  main:
    name: "Quickstart"
    identifier: "tutorials-beginners-quickstart-quickstart"
    parent: "tutorials-top"
    weight: 23
---

This short introduction uses Keras to:

1. Build a neural network that classifies images.
1. Train this neural network.
1. And, finally, evaluate the accuracy of the model.
1. Save and restore the created model.

Before running the quickstart you need to have Keras installed. Please refer to the [installation](/installation/) section for installation instructions.

```{r}
library(keras)
```

Let's start by loading and preparing the [MNIST dataset](http://yann.lecun.com/exdb/mnist/). The values of the pixels are integers between 0 and 255, and we will convert them to floats between 0 and 1.

```{r}
mnist <- dataset_mnist()
mnist$train$x <- mnist$train$x/255
mnist$test$x <- mnist$test$x/255
```

Now, let's define the a Keras model using the sequential API.

```{r}
model <- keras_model_sequential() %>% 
  layer_flatten(input_shape = c(28, 28)) %>% 
  layer_dense(units = 128, activation = "relu") %>% 
  layer_dropout(0.2) %>% 
  layer_dense(10, activation = "softmax")
```

**Note** that when using the Sequential API the first layer must specify the `input_shape` argument which represents the dimensions of the input. In our case, the images are 28x28.

After defining the model, you can see information about layers, number of parameters, etc. with the `summary` function:

```{r}
summary(model)
```

The next step after building the model is to compile it. It's at compile time that we define what **loss** will be optimized and what **optimizer** will be used. You can also specify metrics, callbacks etc. that are meant to be run during the model fitting.

Compiling is done with the `compile` function:

```{r}
model %>% 
  compile(
    loss = "sparse_categorical_crossentropy",
    optimizer = "adam",
    metrics = "accuracy"
  )
```

**Note** that `compile` and `fit` (which we are going to see next) modify the `model` object in place, unlike most R functions. 

Now let's fit our model:

```{r}
model %>% 
  fit(
    x = mnist$train$x, y = mnist$train$y,
    epochs = 5,
    validation_split = 0.3,
    verbose = 2
  )
```

We can now make predictions with our model using the `predict` function:

```{r}
predictions <- predict(model, mnist$test$x)
head(predictions, 2)
```

By default `predict` will return the output of the last Keras layer.
In our case this is the probability for each class. You can also use `predict_classes` and `predict_proba` to generate `class` and  `probability` outputs - these functions are slightly different from `predict`, since they will be run in batches.

You can assess model performance on a different dataset using the `evaluate` function, for example:

```{r}
model %>% 
  evaluate(mnist$test$x, mnist$test$y, verbose = 0)
```

Our model achieved ~90% accuracy on the test set. 

Unlike models built with the `lm` function, to save Keras models for later prediction, you need to use specialized functions, like `save_model_tf`:

```{r}
save_model_tf(object = model, filepath = "model")
```

You can then reload the model and use it to make predictions:

```{r}
reloaded_model <- load_model_tf("model")
all.equal(predict(model, mnist$test$x), predict(reloaded_model, mnist$test$x))
```

More information about saving and serializing models, as well as about different model types, is available in the guides.

```{r include=FALSE}
system("rm -r model")
```

