---
title: "Quick start"
type: docs
menu:
  main:
    name: "Quick start"
    identifier: "installation-get-started"
    parent: "installation-overview"
    weight: 10
aliases:
  - /tensorflow/articles/installation.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Prior to using the **tensorflow** R package you need to install a version of TensorFlow 
on your system. Below we describe how to install TensorFlow as well the various 
options available for customizing your installation. 

Note that this article principally covers the use of the R `install_tensorflow()` function, 
which provides an easy to use wrapper for the various steps required to install 
TensorFlow.

You can also choose to install TensorFlow manually (as described at <https://www.tensorflow.org/install/>). In that case the [Custom Installation](/installation/custom) section 
covers how to arrange for the tensorflow R package to use the version you installed.

TensorFlow is tested and supported on the following 64-bit systems:

- Ubuntu 16.04 or later
- Windows 7 or later
- macOS 10.12.6 (Sierra) or later (no GPU support)

## Installation

First, install the tensorflow R package from GitHub as follows:

```{r, eval = FALSE}
install.packages("tensorflow")
```

Then, use the `install_tensorflow()` function to install TensorFlow. **Note that** on Windows you need a working installation of [Anaconda](https://www.anaconda.com/distribution/).

```{r, eval = FALSE}
library(tensorflow)
install_tensorflow()
```

You can confirm that the installation succeeded with:

```{r, eval = TRUE}
library(tensorflow)
tf$constant("Hellow Tensorflow")
```

This will provide you with a default installation of TensorFlow suitable for use 
with the tensorflow R package. Read on if you want to learn about additional installation options, including installing a version of TensorFlow that takes advantage of Nvidia GPUs if you have the correct CUDA libraries installed.

## Installation methods

TensorFlow is distributed as a Python package and so needs to be installed within a Python environment on your system. By default, the `install_tensorflow()` function attempts to install TensorFlow within an isolated Python environment ("r-reticulate").

These are the available methods and their behavior:

| Method  | Description  |
|--------------|----------------------------------------------------------|
| auto  | Automatically choose an appropriate default for the current platform. |
| virtualenv  | Install into a Python virtual environment at `~/.virtualenvs/r-reticulate` |
| conda  | Install into an Anaconda Python environment named `r-reticulate` |
| system  | Install into the system Python environment  |

The "virtualenv" and "conda" methods are available on Linux and OS X and only the "conda" method is available on Windows. 

`install_tensorflow` is a wraper around `reticulate::py_install`. Please refer to ['Installing Python Packages'](https://rstudio.github.io/reticulate/articles/python_packages.html) for more information.

## Alternate Versions

By default, `install_tensorflow()` install the latest release version of TensorFlow. You can override this behavior by specifying the `version` parameter. For example:

```{r}
install_tensorflow(version = "2.0.0")
```

Note that this should be a full `major.minor.patch` version specification (rather than just major and minor versions).

You can install the nightly build of TensorFlow (CPU or GPU version) with:

```{r}
install_tensorflow(version = "nightly")      # cpu version
install_tensorflow(version = "nightly-gpu")  # gpu version
```

You can install any other build of TensorFlow by specifying a URL to a TensorFlow binary. For example:

```{r}
install_tensorflow(version = "https://files.pythonhosted.org/packages/c2/c1/a035e377cf5a5b90eff27f096448fa5c5a90cbcf13b7eb0673df888f2c2d/tf_nightly-1.12.0.dev20180918-cp36-cp36m-manylinux1_x86_64.whl")
```

  