---
title: "Eager execution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Eager execution} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
type: "docs"
menu:
  main:
    name: "Eager execution"
    identifier: "custom-basic-eager-execution"
    parent: "custom-basic-top"
    weight: 10
aliases:
  - /tensorflow/
  - /tensorflow/articles/basic_usage.html
  - /tensorflow/articles/using_tensorflow_api.html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```


TensorFlow's eager execution is an imperative programming environment that
evaluates operations immediately, without building graphs: operations return
concrete values instead of constructing a computational graph to run later. This
makes it easy to get started with TensorFlow and debug models, and it
reduces boilerplate as well. To follow along with this guide, run the code
samples below in an interactive `R` interpreter.

Eager execution is a flexible machine learning platform for research and
experimentation, providing:

* *An intuitive interface*—Structure your code naturally and use R data
  structures. Quickly iterate on small models and small data.
* *Easier debugging*—Call ops directly to inspect running models and test
  changes. Use standard R debugging tools for immediate error reporting.
* *Natural control flow*—Use R control flow instead of graph control
  flow, simplifying the specification of dynamic models.

Eager execution supports most TensorFlow operations and GPU acceleration.

Note: Some models may experience increased overhead with eager execution
enabled. Performance improvements are ongoing, but please
[file a bug](https://github.com/tensorflow/tensorflow/issues) if you find a
problem and share your benchmarks.

## Setup and basic usage

```{r}
library(tensorflow)
library(tfautograph)
library(keras)
library(tfdatasets)
```

In Tensorflow 2.0, eager execution is enabled by default.

```{r}
tf$executing_eagerly()
```

Now you can run TensorFlow operations and the results will return immediately:

```{r}
x <- matrix(2, ncol = 1, nrow = 1)
m <- tf$matmul(x, x)
m
```

Enabling eager execution changes how TensorFlow operations behave—now they
immediately evaluate and return their values to R `tf$Tensor` objects
reference concrete values instead of symbolic handles to nodes in a computational
graph. Since there isn't a computational graph to build and run later in a
session, it's easy to inspect results using `print()` or a debugger. Evaluating,
printing, and checking tensor values does not break the flow for computing
gradients.

Eager execution works nicely with R. TensorFlow
[math operations](https://www.tensorflow.org/api_guides/python/math_ops) convert
R objects and R arrays to `tf$Tensor` objects. The
`as.array` method returns the object's value as an R `array`.

```{r}
a <- tf$constant(matrix(c(1,2,3,4), ncol = 2))
a
```

```{r}
# Broadcasting support
b <- tf$add(a, 1)
b
```

```{r}
# Operator overloading is supported
a * b
```

```{r}
# Obtain an R value from a Tensor
as.array(a)
```

## Dynamic control flow

A major benefit of eager execution is that all the functionality of the host
language is available while your model is executing. So, for example,
it is easy to write [fizzbuzz](https://en.wikipedia.org/wiki/Fizz_buzz):

```{r}
fizzbuzz <- autograph(function(max_num) {
  counter <- tf$constant(0)
  max_num <- tf$convert_to_tensor(max_num)
  for (num in (tf$range(max_num) + 1)) {
    if ((num %% 3 == 0) & (num %% 5 == 0)) {
      tf$print("FizzBuzz")
    } else if (num %% 3 == 0) {
      tf$print("Fizz")
    } else if (num %% 5 == 0) {
      tf$print("Buzz")
    } else {
      tf$print(num)
    }
    counter <- counter + 1
  }
})
fizzbuzz(15)
```

This has conditionals that depend on tensor values and it prints these values
at runtime.

## Eager training

### Computing gradients

[Automatic differentiation](https://en.wikipedia.org/wiki/Automatic_differentiation)
is useful for implementing machine learning algorithms such as
[backpropagation](https://en.wikipedia.org/wiki/Backpropagation) for training
neural networks. During eager execution, use `tf$GradientTape` to trace
operations for computing gradients later.

You can use `tf$GradientTape` to train and/or compute gradients in eager. It is especially useful for complicated training loops.  

Since different operations can occur during each call, all
forward-pass operations get recorded to a "tape". To compute the gradient, play
the tape backwards and then discard. A particular `tf$GradientTape` can only
compute one gradient; subsequent calls throw a runtime error.

```{r}
w <- tf$Variable(1)
with(tf$GradientTape() %as% tape, {
  loss <- w * w
})
grad <- tape$gradient(loss, w)
grad
```

### Train a model

The following example creates a multi-layer model that classifies the standard
MNIST handwritten digits. It demonstrates the optimizer and layer APIs to build
trainable graphs in an eager execution environment.

```{r}
# Fetch and format the mnist data
mnist <- dataset_mnist()
dataset <- tensor_slices_dataset(mnist$train) %>% 
  dataset_map(function(x) {
    x$x <- tf$cast(x$x, tf$float32)/255
    x$x <- tf$expand_dims(x$x, axis = -1L)
    unname(x)
  }) %>% 
  dataset_shuffle(1000) %>% 
  dataset_batch(32)

dataset
```

```{r}
mnist_model <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 16, kernel_size = c(3,3), activation= "relu",
                input_shape = shape(NULL, NULL, 1)) %>% 
  layer_conv_2d(filters = 16, kernel_size = c(3,3), activation = "relu") %>% 
  layer_global_average_pooling_2d() %>% 
  layer_dense(units = 10)
```

Even without training, call the model and inspect the output in eager execution:

```{r}
el <- dataset %>% 
  dataset_take(1) %>% 
  dataset_collect()
mnist_model(el[[1]])
```

While keras models have a builtin training loop (using the fit method), sometimes you need more customization. Here's an example, of a training loop implemented with eager:

```{r}
optimizer <- optimizer_adam()
loss_object <- tf$keras$losses$SparseCategoricalCrossentropy(from_logits = TRUE)

loss_history <- list()
```

Note: Use the assert functions in `tf$debugging` to check if a condition holds up. This works in eager and graph execution.

```{r}
train_step <- function(images, labels) {
  with(tf$GradientTape() %as% tape, {
    logits <- mnist_model(images, training = TRUE)
    tf$debugging$assert_equal(logits$shape, shape(32, 10))
    loss_value <- loss_object(labels, logits)
  })
  loss_history <<- append(loss_history, loss_value)
  grads <- tape$gradient(loss_value, mnist_model$trainable_variables)
  optimizer$apply_gradients(
    purrr::transpose(list(grads, mnist_model$trainable_variables))
  )
}

train <- autograph(function() {
  for (epoch in seq_len(3)) {
    for (batch in dataset) {
      train_step(batch[[1]], batch[[2]])
    }
    tf$print("Epoch", epoch, "finished.")
  }
})

train()
```

```{r}
history <- loss_history %>% 
  purrr::map(as.numeric) %>% 
  purrr::flatten_dbl()
ggplot2::qplot(x = seq_along(history), y = history, geom = "line")
```

### Variables and optimizers

`tf$Variable` objects store mutable `tf$Tensor`-like values accessed during
training to make automatic differentiation easier. 

The collections of variables can be encapsulated into layers or models, along with methods that operate on them. See [Custom Keras layers and models](../../keras/custom_layers/) for details. The main difference between layers and models is that models add methods like  `fit`, `evaluate`, and `save`.

For example, the automatic differentiation example above
can be rewritten:

```{r}
Linear <- function() {
  keras_model_custom(model_fn = function(self) {
    self$w <- tf$Variable(5, name = "weight")
    self$b <- tf$Variable(10, name = "bias")
    function(inputs, mask = NULL, training = TRUE) {
      inputs*self$w + self$b
    }
  }) 
}
```

```{r}
# A toy dataset of points around 3 * x + 2
NUM_EXAMPLES <- 2000
training_inputs <- tf$random$normal(shape = shape(NUM_EXAMPLES))
noise <- tf$random$normal(shape = shape(NUM_EXAMPLES))
training_outputs <- training_inputs * 3 + 2 + noise

# The loss function to be optimized
loss <- function(model, inputs, targets) {
  error <- model(inputs) - targets
  tf$reduce_mean(tf$square(error))
}

grad <- function(model, inputs, targets) {
  with(tf$GradientTape() %as% tape, {
    loss_value <- loss(model, inputs, targets)
  })
  tape$gradient(loss_value, list(model$w, model$b))
}
```

Next:

1. Create the model.
2. The Derivatives of a loss function with respect to model parameters.
3. A strategy for updating the variables based on the derivatives.

```{r}
model <- Linear()
optimizer <- optimizer_sgd(lr = 0.01)

cat("Initial loss: ", as.numeric(loss(model, training_inputs, training_outputs), "\n"))

for (i in seq_len(300)) {
  grads <- grad(model, training_inputs, training_outputs)
  optimizer$apply_gradients(purrr::transpose(
    list(grads, list(model$w, model$b))
  ))
  if (i %% 20 == 0)
    cat("Loss at step ", i, ": ", as.numeric(loss(model, training_inputs, training_outputs)), "\n")
}

```

```{r}
model$w
model$b
```

Note: Variables persist until the last reference to the object
is removed, and is the variable is deleted.

### Object-based saving

A Keras model includes a convinient `save_weights` method allowing you to easily create a checkpoint: 

```{r}
save_model_weights_tf(model, "weights")
load_model_weights_tf(model, filepath = "weights")
```

Using `tf$train$Checkpoint` you can take full control over this process.

This section is an abbreviated version of the [guide to training checkpoints](./checkpoint/).

```{r}
x <- tf$Variable(10)
checkpoint <- tf$train$Checkpoint(x = x)
```

```{r}
x$assign(2) # Assign a new value to the variables and save.
checkpoint_path <- "ckpt/"
checkpoint$save(checkpoint_path)
```

```{r}
x$assign(11) # Change the variable after saving.
checkpoint$restore(tf$train$latest_checkpoint(checkpoint_path))
x
```

To save and load models, `tf$train$Checkpoint` stores the internal state of objects,
without requiring hidden variables. To record the state of a `model`,
an `optimizer`, and a global step, pass them to a `tf$train$Checkpoint`:

```{r}
model <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 16, kernel_size = c(3,3), activation = "relu") %>% 
  layer_global_average_pooling_2d() %>% 
  layer_dense(units = 10)

optimizer <- optimizer_adam(lr = 0.001)

checkpoint_dir <- 'path/to/model_dir'
if (!dir.exists(checkpoint_dir))
  dir.create(checkpoint_dir, recursive = TRUE)

checkpoint_prefix <- file.path(checkpoint_dir, "ckpt")

root <- tf$train$Checkpoint(optimizer = optimizer, model = model)

root$save(checkpoint_prefix)
root$restore(tf$train$latest_checkpoint(checkpoint_dir))
```

Note: In many training loops, variables are created after tf$train$Checkpoint.restore is called. These variables will be restored as soon as they are created, and assertions are available to ensure that a checkpoint has been fully loaded. See the guide to training checkpoints for details.

### Object-oriented metrics

`tf$keras$metrics` are stored as objects. Update a metric by passing the new data to
the callable, and retrieve the result using the `tf$keras$metrics$result` method,
for example:

```{r}
m <- tf$keras$metrics$Mean("loss")
m(0)
m(5)
m$result()
m(c(8, 9))
m$result()
```

### Summaries and TensorBoard

[TensorBoard](https://tensorflow.org/tensorboard) is a visualization tool for
understanding, debugging and optimizing the model training process. It uses
summary events that are written while executing the program.

You can use `tf$summary` to record summaries of variable in eager execution.
For example, to record summaries of `loss` once every 100 training steps:

```{r, eval = FALSE}
logdir <- "./tb/"
writer = tf$summary$create_file_writer(logdir)
tensorboard(log_dir = logdir) # This will open a browser window pointing to Tensorboard

with(writer$as_default(), {
  for (step in seq_len(1000)) {
    # Calculate loss with your real train function.
    loss = 1 - 0.001 * step
    if (step %% 100 == 0)
      tf$summary$scalar('loss', loss, step=step)
  }
})
```

## Advanced automatic differentiation topics

### Dynamic models

`tf$GradientTape` can also be used in dynamic models. This example for a
[backtracking line search](https://wikipedia.org/wiki/Backtracking_line_search)
algorithm looks like normal R code, except there are gradients and is
differentiable, despite the complex control flow:

```{r}
line_search_step <- tf$custom_gradient(autograph(function(fn, init_x, rate = 1) {
  with(tf$GradientTape() %as% tape, {
    tape$watch(init_x)
    value <- fn(init_x)
  })
  grad <- tape$gradient(value, init_x)
  grad_norm <- tf$reduce_sum(grad * grad)
  init_value <- value
  while(value > (init_value - rate * grad_norm)) {
    x <- init_x - rate * grad
    value <- fn(x)
    rate = rate/2
  }
  list(x, value)
}))
```

### Custom gradients

Custom gradients are an easy way to override gradients. Within the forward function, define the gradient with respect to the
inputs, outputs, or intermediate results. For example, here's an easy way to clip
the norm of the gradients in the backward pass:

```{r}
clip_gradient_by_norm <- function(x, norm) {
  y <- tf$identity(x)
  grad_fn <- function(dresult) {
    list(tf$clip_by_norm(dresult, norm), NULL)
  }
  list(y, grad_fn)
}
```

Custom gradients are commonly used to provide a numerically stable gradient for a
sequence of operations:

```{r}
log1pexp <- function(x) {
  tf$math$log(1 + tf$exp(x))
}

grad_log1pexp <- function(x) {
  with(tf$GradientTape() %as% tape, {
    tape$watch(x)
    value <- log1pexp(x)
  })
  tape$gradient(value, x)
}
```

```{r}
# The gradient computation works fine at x = 0.
grad_log1pexp(tf$constant(0))
```
```{r}
# However, x = 100 fails because of numerical instability.
grad_log1pexp(tf$constant(100))
```

Here, the `log1pexp` function can be analytically simplified with a custom
gradient. The implementation below reuses the value for `tf$exp(x)` that is
computed during the forward pass—making it more efficient by eliminating
redundant calculations:

```{r}
log1pexp <- tf$custom_gradient(f = function(x) {
  e <- tf$exp(x)
  grad_fn <- function(dy) {
    dy * (1 - 1/(e + e))
  }
  list(tf$math$log(1 + e), grad_fn)
})

grad_log1pexp <- function(x) {
  with(tf$GradientTape() %as% tape, {
    tape$watch(x)
    value <- log1pexp(x)
  })
  tape$gradient(value, x)
}
```

```{r}
# As before, the gradient computation works fine at x = 0.
grad_log1pexp(tf$constant(0))
```

```{r}
# And the gradient computation also works at x = 100.
grad_log1pexp(tf$constant(100))
```

## Performance

Computation is automatically offloaded to GPUs during eager execution. If you
want control over where a computation runs you can enclose it in a
`tf$device('/gpu:0')` block (or the CPU equivalent):

```{r, eval = FALSE}
fun <- function(device, steps = 200) {
  with(tf$device(device), {
    x <- tf$random$normal(shape = shape)
    for (i in seq_len(steps)) {
      tf$matmul(x, x)  
    }  
  })
}
microbenchmark::microbenchmark(
  fun("/cpu:0"),
  fun("/gpu:0")
)
```

```{r}
# Unit: milliseconds
#           expr      min        lq      mean    median        uq       max neval
#  fun("/cpu:0") 1117.596 1135.5450 1165.6269 1157.2208 1195.1529 1300.2236   100
#  fun("/gpu:0")  112.888  121.7164  127.8525  126.6708  132.0415  228.1009   100
```

A `tf$Tensor` object can be copied to a different device to execute its
operations:

```{r, eval = FALSE}
x <- tf$random$normal(shape = shape(10,10))

x_gpu0 <- x$gpu()
x_cpu <- x$cpu()

tf$matmul(x_cpu, x_cpu)    # Runs on CPU
tf$matmul(x_gpu0, x_gpu0)  # Runs on GPU:0
```

### Benchmarks

For compute-heavy models, such as
[ResNet50](https://github.com/tensorflow/tensorflow/tree/master/tensorflow/contrib/eager/python/examples/resnet50)
training on a GPU, eager execution performance is comparable to `tf_function` execution.
But this gap grows larger for models with less computation and there is work to
be done for optimizing hot code paths for models with lots of small operations.

## Work with functions

While eager execution makes development and debugging more interactive,
TensorFlow 1.x style graph execution has advantages for distributed training, performance
optimizations, and production deployment. To bridge this gap, TensorFlow 2.0 introduces `function`s via the `tf_function` API. For more information, see the [tf_function](./function) guide.



